---
title: "Machine Learning final assignment"
author: "Hatsumi Hirota"
date: "July 17, 2016"
output: html_document
---

##Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

##Objective
The goal of this project is to predict the manner in which they did the exercise. In the end, I will also use your prediction model to predict 20 different test cases.

##Data Source
The training data for this project are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

##Get necessary packages
```{r}
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(randomForest)
```

##Read data into R
```{r}
setwd("~/Documents/coursera/8. Practical Machine Learning")
training = read.csv("pml-training.csv", na.strings = c("NA", ""))
testing = read.csv("pml-testing.csv", na.strings = c("NA", ""))

dim(training)
dim(testing)
```

##Cleaning Data
```{r}
##Remove nearzerovariance variables
nzv <- nearZeroVar(training, saveMetrics=TRUE)
training <- training[,nzv$nzv==FALSE]

##Remove the first column
training <- training[c(-1)]

##Remove variables with full of NA and make testing dataset consistent with training dataset in terms of available columns
features <- names(training[,colSums(is.na(training)) == 0])[1:57]
training<-training[,c(features,"classe")]
testing<-testing[,c(features)]

for (i in 1:length(testing) ) {
  for(j in 1:length(training)) {
    if( length( grep(names(training[i]), names(testing)[j]) ) ==1)  {
      class(testing[j]) <- class(training[i])
    }      
  }      
}

testing <- rbind(training[2, -58] , testing) 
testing <- testing[-1,]
```

##Partitioning the training sets 60% for mytraining and 40% for mytesting
```{r}
inTrain <- createDataPartition(y=training$classe, p=0.6, list=FALSE)
mytraining <- training[inTrain, ]; mytesting <- training[-inTrain, ]
dim(mytraining)
dim(mytesting)
```

##Decision Tree
Decision Tree generated the accuracy level of 0.8703
```{r}
DT <- rpart(classe ~ ., data=mytraining, method="class")
fancyRpartPlot(DT)
DT2 <- predict(DT, mytesting, type = "class")
confusionMatrix(DT2, mytesting$classe)
```

##Ramdom Forests
Random Forests generated the accuracy of 0.9981, so this is better approach than Decision Tree
```{r}
RF <- randomForest(classe ~. , data=mytraining)
RF2 <- predict(RF, mytesting, type = "class")
confusionMatrix(RF2, mytesting$classe)
```


##Out of sample error
Since Random Forests method was more accurate, I will use this to predict 20 cases from testing dataset
```{r}
predictions <- predict(RF, testing, type = "class")
```

